\mbox{Verifying Imperative Higher-Order Programs}

### 2009-10-07 papers/esop2010/main.tex
go
section headings

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc} 
\input{utf8symbols}
\usepackage{stmaryrd}
\usepackage{amsmath,amssymb}
\usepackage{mathpartir}

@@

%external TODO : proofs of system
%

@@

\paragraph{Correctness}

In the 
%TODO lemmas about correctness
\section{Examples}
\label{sec:examples}

@@

>>>>

\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc} 
\input{utf8symbols}
\usepackage{stmaryrd}
\usepackage{amsmath,amssymb}
\usepackage{mathpartir}
\usepackage{url}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}

@@

@@

\subsection{Correctness}

We have proved the correctness of our system. All the properties are proved in
subject reduction style, maintaining an invariant during the execution of the
program, modeled with small-step semantics. We only consider well-typed terms.

For the \wpre-calculus, we want to obtain that if we can prove $\wpre_s(t,q)$
where $s$ is the current state of the store, $t$ a program, and $q$ the
postcondition, and if $t$ reduces to the (syntactic) value $v$ and the store
$s'$, then $q~s'~v$ is valid. The subject reduction way of proving this
proceeds by proving that the validity of the formula $\wpre_s(t,q)$ is in fact
maintained during the execution of the program. However, in the presence of
effect typing, the effect of $t$ may {\em decrease} during program execution.
Thus, the main lemma for the \wpre-calculus is 
\begin{lemma} 
  $\wpre_s(t,q)$ and $s,t ->> s',t'$ implies $\wpre_{s'}( (t : ε'<:ε), q)$.
\end{lemma}
Here, $->>$ means ``reduces in zero or more steps to''. The main theorem
is then an easy consequence.

For the alias restriction, or more precisely the restriction on reference
creation, we formally prove that each region contains exactly one memory
location, i.e. is actually a {\em singleton region}. Here we have to
instrument the semantics and mark each reference cell in the store with the
region it belongs to; therefore the exact statement is the following:
\begin{theorem}
  If $s,t ->>s',t$ and if $s$ contains only singleton regions (each reference
  cell maps to a different region), then this is also the case for $s'$.
\end{theorem}

Both theorems are proved in the technical appendix~\cite{appendixesop2010}
accompanying this paper. The aliasing restriction is proved there in a more
general setting, while the correctness of the \wpre-calculus is proved in a
slightly different version.

@@

\section{Examples}
\label{sec:examples}
